Corotating Dust Clumps Around Adolescent M-Dwarfs: Four Years of TESS

### 2023-06-26 paper/ms.tex
update ms

Of our \ncpvsfound\ confirmed and candidate CQVs, \nnotfieldbanyan\
were confidently associated with a nearby moving group or open cluster
using BANYAN\,$\Sigma$.  The relevant groups are listed in
Table~\ref{tab:thetable}; their ages span $\approx$5-150\,Myr.  The
most prodigious groups were Sco-Cen, Tuc-Hor, and Columba.  The yield
in Sco-Cen (LCC / UCL / USco) is not surprising, since Sco-Cen
contains the majority of pre-main-sequence stars in the solar
neighborhood.  The absence of CQVs in IC\,2602 and IC\,2391 can be
understand by those clusters being beyond 100\,pc, where the
completeness of the TESS 120-second cadence data is below 10\%; based
on the abundance of CQVs in the 40\,Myr Tuc-Hor and Columba, the
similarly-aged IC\,2602 and IC\,2391 clusters should have a large
number of CQVs in the TESS full frame image data.

Of the \ngoodsfieldbanyan\ stars for which BANYAN\,$\Sigma$ did not
find any association, one (TIC~302160226) is a member of $\alpha$~Per
($t\approx 86\pm16$\,Myr;
\citealt{2021A&A...645A..84M,2022arXiv221109822B}).  For the other two
(TIC~58084670 and TIC~141146667), we are convinced of both their CQV
status and their youth, but we were not able to confidently associate
either star with any young groups.  Both objects were noted as being in the
``diffuse'' Class~II YSO population near the Sun by \citet{2021ApJ...917...23K}.
That diffuse population may contain kinematic structures that have yet
to be unravelled.  However, the absence of obvious young, co-moving
stars within $\Delta r$=$40$\,pc and $\Delta v$=$3$\,km\,s$^{-1}$ of
our two CQVs suggest that they may have ``run away'' from their birth
locations.

@@

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{tic4029_segments.pdf}
		\vspace{-0.2cm}
		\caption{
      {\bf LP 12-502 light curve}, where each time chunk represents one TESS orbit.
      Flares are drawn in gray.
      The red vertical lines highlight apparently instantaneous state changes in
      the shape of the dip pattern.
      The light curve is binned to 15-minute intervals so that
      there are 96 points per day.  Data gaps are not drawn.
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.43\textwidth]{TIC402980664_river_gist_stern_0_64_manual_20230617_mask_v0_nterms2_vmin-0.04000_vmax0.01000.png}
			\includegraphics[width=0.43\textwidth]{ANNOTATED_TIC402980664_river_gist_stern_248_315_manual_20230617_mask_v0_nterms2_vmin-0.04000_vmax0.01000.png}
		}
	\vspace{-0.1cm}
			\subfloat{
		\includegraphics[width=0.43\textwidth]{ANNOTATED_TIC402980664_river_gist_stern_1411_1481_manual_20230617_mask_v0_nterms2_vmin-0.04000_vmax0.01000.png}
				\includegraphics[width=0.43\textwidth]{ANNOTATED_TIC402980664_river_gist_stern_1233_1265_manual_20230617_mask_v0_nterms2_vmin-0.04000_vmax0.01000.png}
	}
	\end{center}
	\caption{
    {\bf River plots of LP 12-502}, showing (clockwise from top-left)
    Sectors 18-19, 25-26, 53, and 58-59.  A two-harmonic sinusoid has
    been subtracted over specific chunks in time ({\bf see text}).
    For Sectors 25-26 (cycles 248-315), three periods are overplotted:
    $P$=18.5611\,hr (gray vertical line); 18.5404\,hr (orange); 18.5683\,hr (green).
    For Sector 53, gray is identical, while cyan is 18.5145\,hr.
    For Sectors 58-59, the magenta line is 18.5473\,hr, and the green
    line is 18.5672\,hr.
	}
	\label{fig:lpriver0}
\end{figure*}

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{TIC4029_zoom.png}
		\vspace{-0.45cm}
		\caption{
			{\bf LP 12-502 cycles 1233-1247}, showing the most dramatic state
			change observed over the three year baseline.
		}
		\label{fig:lpzoom}
	\end{center}
\end{figure*}

>>>>

Of our \ncpvsfound\ confirmed and candidate CQVs, \nnotfieldbanyan\
were confidently associated with a nearby moving group or open cluster
using BANYAN\,$\Sigma$.  The relevant groups are listed in
Table~\ref{tab:thetable}; their ages span $\approx$5-150\,Myr.  The
most prodigious groups were Sco-Cen, Tuc-Hor, and Columba.  The yield
in Sco-Cen (LCC / UCL / USco) is not surprising, since Sco-Cen
contains the majority of pre-main-sequence stars in the solar
neighborhood.  The absence of CQVs in IC\,2602 and IC\,2391 can be
understand by those clusters being beyond 100\,pc, where the
completeness of the TESS 120-second cadence data is below 10\%; based
on the abundance of CQVs in the 40\,Myr Tuc-Hor and Columba, the
similarly-aged IC\,2602 and IC\,2391 clusters should show a large
number of CQVs in the TESS full frame image data.

Of the \ngoodsfieldbanyan\ stars for which BANYAN\,$\Sigma$ did not
find any association, one (TIC~302160226) is a member of $\alpha$~Per
($t\approx 86\pm16$\,Myr;
\citealt{2021A&A...645A..84M,2022arXiv221109822B}).  For the other two
(TIC~58084670 and TIC~141146667), we are convinced of both their CQV
status and their youth, but we were not able to confidently associate
either star with any young groups.  Both objects were noted as being in the
``diffuse'' Class~III YSO population near the Sun by \citet{2021ApJ...917...23K}.
That diffuse population may contain kinematic structures that have yet
to be unravelled.  However, the absence of obvious young, co-moving
stars within $\Delta r$=$40$\,pc and $\Delta v$=$3$\,km\,s$^{-1}$ of
our two CQVs suggest that they may have ``run away'' from their birth
locations.

@@

\subsection{Infrared excesses}

As is typical for CQVs \citep{2017AJ....153..152S}, most CQVs in our
catalog did not show infrared excesseses in the WISE1-4 bands.
Visually inspecting the SEDs, we labelled two objects as having
``good'' infrared excesses (both W3 and W4), and three as ``possible''
infrared excess.

The two ``good'' IR excesses were TIC~193136669 (TWA~34) and
TIC~57830249 (TWA~33).  Both are in the TW Hydrae association
($\approx$10\,Myr), and have ``long'' periods of 38\,hr and 44\,hr
respectively.  In our initial labelling, we labelled both as ``maybe''
CQVs because the dips in their light curves did not show the rigid
periodicity typical of CQVs;  their periods were also relatively
long.  After labelling them, we learned that TIC 193136669 has a cold
molecular disk based on observed 1.3\,mm continuum emission and
resolved Keplerian $^{12}{\rm CO}(2-1)$ emission
\citep{2015A&A...582L...5R}.  It has also been labelled a dipper by
\citet{2022ApJS..263...14C}; we agree with their designation, but
nonetheless leave it in our catalog as an indication of possible
ambiguities in our search process.  TIC 57830249 (TWA~33) also has
previously detected 1.3\,mm continuum emission
\citep{2015A&A...582L...5R}, suggestive of cold dust grains being
present.  It therefore probably also a dipper that snuck its way into
our ``maybe'' pile.

Our three ``possible'' infrared excesses were TIC~405910546,
289840926, and 244161191.  After a literature search, we concluded
that none have clear evidence for the presence of a disk.
TIC~405910546, in LCC, shows a unique TESS light curve, reminescent of
a $P$=38\,hr singly-eclipsing binary, except with additional
substructure during each eclipse that resembles the CQV phenomenon
more than any other variability of which we are aware.  TIC~289840926
($\beta$~Pic moving group, $P$=4.8\,hr), show what we believe is a
clear CQV signals, but has no definitive evidence for a large, dusty
disk.  TIC~244161191 (hilariously, TOI-278), in Columba, also has no
definitive evidence for a large disk.  It is however
``multi-periodic''---in addition to the 7.17\,hr CPV signal, there is
a superposed 8.39\,hr sinusoidal signal, probably from an unresolved
binary neighbor.

Comparing these results against the backdrop of our increasing
understanding of dippers
\citep[e.g.][]{2014AJ....147...82C,2016ApJ...816...69A,2021ApJ...908...16R,2022ApJS..263...14C},
it is clear that the loss of an infrared excess is associated with
strong changes in a star's optical variability.  It is not
unreasonable to imagine connections between CQVs and dippers: both
classes of object can show transient flux dips that are relatively
narrow in duration.  Such dips are probably caused by clumps of dust.
However the CQV dips are typically {\it i}) more periodic and {\it
ii}) less deep than those of dippers, and {\it iii}) they display far
less transience over timescales of a few to tens of cycles.  At a
population level, the CQV stars are also older.  A shared mystery
between the CQVs and dippers, in our own estimation however, is the
exact mechanism that produces the {\it narrowness} of the dust clumps
in certain cases.  It is not unreasonable to imagine a similar
mechanism operating for both types of object, tied perhaps to a shared
magnetic topology, or perhaps to a preference for dust to inspiral to
the star in clumps from tidally disrupted infalling bodies.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{tic4029_segments.pdf}
		\vspace{-0.2cm}
		\caption{
      {\bf LP 12-502 light curve}, where each time chunk represents one TESS orbit.
      Flares are drawn in gray.
      The red vertical lines highlight apparently instantaneous state changes in
      the shape of the dip pattern.
      The light curve is binned to 15-minute intervals so that
      there are 96 points per day.  Data gaps with more than one missing 15-minute
      cadence appear in white.
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.43\textwidth]{TIC402980664_river_gist_stern_0_64_manual_20230617_mask_v0_nterms2_vmin-0.04000_vmax0.01000.png}
			\includegraphics[width=0.43\textwidth]{ANNOTATED_TIC402980664_river_gist_stern_248_315_manual_20230617_mask_v0_nterms2_vmin-0.04000_vmax0.01000.png}
		}
	\vspace{-1cm}
			\subfloat{
		\includegraphics[width=0.43\textwidth]{ANNOTATED_TIC402980664_river_gist_stern_1411_1481_manual_20230617_mask_v0_nterms2_vmin-0.04000_vmax0.01000.png}
				\includegraphics[width=0.43\textwidth]{ANNOTATED_TIC402980664_river_gist_stern_1233_1265_manual_20230617_mask_v0_nterms2_vmin-0.04000_vmax0.01000.png}
	}
	\end{center}
	\vspace{-0.4cm}
	\caption{
    {\bf River plots of LP 12-502}, showing (clockwise from top-left)
    Sectors 18-19, 25-26, 53, and 58-59.  A two-harmonic sinusoid has
    been subtracted over specific chunks in time ({\bf see text}).
    For Sectors 25-26 (cycles 248-315), three periods are overplotted:
    $P$=18.5611\,hr (gray vertical line); 18.5404\,hr (orange); 18.5683\,hr (green).
    For Sector 53, gray is identical, while cyan is 18.5145\,hr.
    For Sectors 58-59, the magenta line is 18.5473\,hr, and the green
    line is 18.5672\,hr.
	}
	\label{fig:lpriver0}
\end{figure*}

%
%\begin{figure*}[!t]
%	\begin{center}
%		\centering
%		\includegraphics[width=0.96\textwidth]{TIC4029_zoom.png}
%		\vspace{-0.45cm}
%		\caption{
%			{\bf LP 12-502 cycles 1233-1247}, showing the most dramatic state
%			change observed over the three year baseline.
%		}
%		\label{fig:lpzoom}
%	\end{center}
%\end{figure*}
